Dissertation reading notes

### 2023-07-22 Texfiles/Draft1.tex
Reorganisation of R

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[authoryear]{natbib}
\usepackage{enumerate}
\setlength{\parindent}{0pt}
\usepackage[parfill]{parskip}
\date{Lent 2023}
\title{Dissertation draft}
\author{Tobias Leigh-Wood}
\usepackage[letterpaper, margin=1.1in]{geometry}
\linespread{1.1}

@@

Standard economic theory suggests that retirement annuities should be highly prized by individuals as a way to
insure against the risk of late death \cite{yaari_65}. However, in developed countries rates of annuitization are far below the
levels that theory predicts. In this paper I test two competing hypotheses for the annuity problem: bequests and
pessimistic life expectancy. In most developed countries, individuals are forced to contribute to either a defined
benefit or defined contribution pension. A defined benefit pension accumulates throughout your job tenure as
Under the coalition government in the UK the law regarding the use of private defined
contribution pensions changed. Individuals were no longer forced to annuitise their pension pots and could access
them in a variety of ways such as lump sum withdrawals or income drawdown and subsequently the number of annuities
sold in the UK dropped precipitously.

Depending on the reason for the lack of annuitization in the UK, the consumption response of retirees to the pension
reform will differ. If individuals do not annuitise because of pessimistic life expectancy I will show that their
consumption should increase. If, on the other hand, individuals do not annuitise because of a bequest motive, consumption
should not change much as a result of the reform. I will solve lifecycle models for both of these cases and simulate
consumption decisions with, and without, forced annuitization. I will then use a variety of empirical models to measure
the consumption change in early retirement that resulted from the policy reform. The size and magnitude of this change
will be indicative of the mechanism causing the annuitization problem.

@@

The so called "pensions freedom act" received Royal Asset in December 2014 marked the end of a series
of pension reforms carried out by the coalition government between 2010 and 2015. The reform was announced
in the Spring budget and made it possible to withdraw money from a private pension pot subject to the
marginal rate of income tax that an individual faced. In the June 2010 budget the government made a first
reform to the annuitization rules, creating an minimum income requirement above which individuals would not need to
annuitise more \cite{finance_act_hmt_2011}. However, this was set at £20,000 and therefore few individuals
were eligible. The minimum income requirement was scrapped in the 2014 bill finally eliminating the
compulsory annuities market. The impact of the reforms on annuity demand has been documented by \cite{cannon_et_al_nier_2016}.
Using data from the Association of British Insurers they show that annuity demand dropped by 75\% from
its maximum.

@@

\section{Data}
The main data set I use is the English Longitudinal Study of Ageing (ELSA) \cite{main_elsa_citation}. ELSA repeatedly interviews
individuals over the age of 50 and asks them a range of questions relating to their income and wealth as well as expectations about
the future. Importantly it also includes detailed information pensions including the type of pension that an individual holds when
working, thus I am able to distinguish between individuals who have defined benefit and defined contribution pensions. For ease of access
I use Harmonized ELSA \footnote{"This analysis uses data or information from the Harmonized ELSA dataset and Codebook, Version G.2 as of
  July 2021 developed by the Gateway to Global Aging Data. The development of the Harmonized ELSA was funded by the National
  Institute on Aging (R01 AG030153, RC2 AG036619, R03 AG043052). For more information,
  please refer to https://g2aging.org/.”} which ensures that variables are comparable across waves. Since this only includes a
subset of the questions in ELSA I also supplement it with variables taken directly from the data.

@@

To illustrate the effect the reform had on sales of annuities in the UK I obtained product data from the Financial Conduct Authority. 
These track the sale of different financial products overtime including data on annuity sales.
% Plot showing the drop of in annuity sales

@@

\section{Methodology}

I first solve a modified retirement lifecycle model. The problem that retirees face is as follows. 
Every period retirees solve:
\begin{equation*}
  V_{t}(a_{t}, y) = \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V_{t+1}(a_{t+1}, y) \}
\end{equation*}
subject to their budget constraint
\begin{equation*}
  c_{t} =a_{t}(1 +r) -  a_{t+1} + y
\end{equation*}
where $a_{t}$ are asset holdings in time t, $y$ is constant income for all periods and . Income can come either
from state pensions, defined benefit pension plans or purchased annuities.
\begin{equation*}
  u(c_{t}) = \frac{c_{t}^{1 - \sigma}}{1 - \sigma}
\end{equation*} In some specifications retirees can leave bequests, I use the bequest function from
\cite{lockwood_red_2012}.
\begin{equation*}
  B(a_{t}) = \bigl( \frac{m}{1 - m} \bigr)^{\sigma}  \frac{(\frac{m}{1 - m}c_{0} + a_{t})^{(1 - \sigma)}}{1 - \sigma}
\end{equation*}
Where $b_{t}$ is the amount left at death, $m$ is a measure of bequest motive strength and $c_{0}$ is
a minimum amount of consumption that individuals want. \textbf{Check this}

@@

The ta

>>>>

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[authoryear]{natbib}
\usepackage{enumerate}
\usepackage{booktabs}
\setlength{\parindent}{0pt}
\usepackage[parfill]{parskip}
\date{Lent 2023}
\title{Dissertation draft}
\author{Tobias Leigh-Wood}
\usepackage[letterpaper, margin=1.1in]{geometry}
\linespread{1.1}

@@

Standard economic theory suggests that retirement annuities should be highly prized by individuals as a way to
insure against the risk of late death \cite{yaari_65}. However, in developed countries rates of annuitization are far below the
levels that theory predicts.
Under the coalition government in the UK the law regarding the use of private defined
contribution pensions changed. Individuals were no longer forced to annuitise their pension pots and could access
them in a variety of ways such as a lump sum withdrawal on retirement or income drawdown and subsequently the number of annuities
sold in the UK dropped precipitously.

I first use the policy reform as a discontintuity to measure the impact of forced annutisation on the consumption
of individuals in the first few years of retirement. Given that the reform was implemented suddenly and without
advanced notice before the Spring 2014 budget I claim that individuals with a retirement year in 2015, 2016 or 2017
are otherwise similar to those with a retirement year in 2011, 2012 and 2013 but the do not need to annuitise their
defined contribution pension pots.

In this paper I test two competing hypotheses for the annuity problem: bequests and
pessimistic life expectancy. Depending on the reason for the lack of annuitization in the UK, the consumption response of retirees to the pension
reform will differ. If individuals do not annuitise because of pessimistic life expectancy I will show that their
consumption should increase. If, on the other hand, individuals do not annuitise because of a bequest motive, consumption
should not change much as a result of the reform. I will solve lifecycle models for both of these cases and simulate
consumption decisions with, and without, forced annuitization. I will then use a variety of empirical models to measure
the consumption change in early retirement that resulted from the policy reform. The size and magnitude of this change
will be indicative of the mechanism causing the annuitization problem.

@@

@@

\section{Data and Policy reform}

\subsection{Data}

The main data set I use is the English Longitudinal Study of Ageing (ELSA) \cite{main_elsa_citation}. ELSA repeatedly interviews
individuals over the age of 50 and asks them a range of questions relating to their income and wealth as well as expectations about
the future. Importantly it also includes detailed information pensions including the type of pension that an individual holds when
working, thus I am able to distinguish between individuals who have defined benefit and defined contribution pensions. For ease of access
I use Harmonized ELSA \footnote{"This analysis uses data or information from the Harmonized ELSA dataset and Codebook, Version G.2 as of
  July 2021 developed by the Gateway to Global Aging Data. The development of the Harmonized ELSA was funded by the National
  Institute on Aging (R01 AG030153, RC2 AG036619, R03 AG043052). For more information,
  please refer to https://g2aging.org/.”} which ensures that variables are comparable across waves. Since this only includes a
subset of the questions in ELSA I also supplement it with variables taken directly from the data.

@@

To illustrate the effect the reform had on sales of annuities in the UK I obtained product data from the Financial Conduct Authority.
These track the sale of different financial products overtime including data on annuity sales.

% Plot showing the drop of in annuity sales

@@

\input{tables/r_summary_stats.tex}

\subsection{Policy reform}

\textbf{Add some other background to pensions in the UK. What does the Blundell etc have
  to say}

Successive governments in the UK have attempted to reform both the public and private pension system. Prior
to 1987 participation in private schemes was limited to employees of firms that had offered one, and there
were few alternatives to the public state pension or a define benefit scheme that a public sector employee
would offer. \textbf{Were there pension reforms in 1990s and 00s}

Announced in the Spring budge of 2014, the so called "pensions freedom act" received Royal Asset in December 2014 marked the end of a series
of pension reforms carried out by the coalition government between 2010 and 2015. The reform was announced
in the Spring budget and made it possible to withdraw money from a private pension pot subject to the
marginal rate of income tax that an individual faced. In the June 2010 budget the government made a first
reform to the annuitization rules, creating an minimum income requirement above which individuals would not need to
annuitise more \cite{finance_act_hmt_2011}. However, this was set at £20,000 and therefore few individuals
were eligible. The minimum income requirement was scrapped in the 2014 bill finally eliminating the
compulsory annuities market. The impact of the reforms on annuity demand has been documented by \cite{cannon_et_al_nier_2016}.
Using data from the Association of British Insurers they show that annuity demand dropped by 75\% from
its maximum.

After the announcement in spring 2014 sales of annuities dropped massively and income drawdown products increased.
However, pension pots could also be taken out in one go and this became a much more common option for many people.

Figure \ref{fig:annovertime} shows purchases of annuities overtime and demonstrates the sharp decrease
in purchases that happened from 2014 to 2015. Likewise there was an increase in the number of pensions that were
being accessed using an income drawdown product, but these do not fully account for the drop in annuitisation. So,
either they are withdrawn at once or they are left alone.

Figure \ref{fig:ann2122} shows the current distribution of how pension pots are first accessed at retirement.
In 2021, 196,736 pots were fully withdrawn at retirement, accounting for over 50\% of pots, prior to the policy
reform this was not the case and most defined contribution pensions were accessed through annuities.

\begin{figure}[h]
  \caption{Pension pots accessed}
  \centering
  \includegraphics[width=0.7\columnwidth]{figures/annuity_overtime.pdf}
  \label{fig:annovertime}
\end{figure}

\begin{figure}[h]
  \caption{How pension pots are accessed}
  \centering
  \includegraphics[width=0.7\columnwidth]{figures/annuity_pot_sizes.pdf}
  \label{fig:ann2122}
\end{figure}

\subsection{Covariate Balance}

I regress demographic and financial characteristics of individuals on year of birth and the treatment dummy.
We can then see if the treatment groups differ on key characteristics such as financial wealth or retirement age.
For a regression discontinuity to be valid we need the treatment and control groups to be similar along
all other characteristics apart from the treatment variable.

In particular I run:
\begin{equation*}
  Y_{i} = \alpha + \beta Post_{i} + \gamma YOB + \kappa (Post_{i} \times YOB_{i}) + \epsilon_{i}
\end{equation*}

\input{tables/cov_balance.tex}
Table \ref{tab:cov_balance} shows these results. We can see ...

Another threat to validity is manipulation of the running variable. This is probably quite a threat in our case
as individuals could delay retirement and/or the purchase of an annuity.

\section{Methodology and Econometric Strategy}

I first solve a modified retirement lifecycle model. The problem that retirees face is as follows.
Every period retirees solve:
\begin{equation*}
  V_{t}(a_{t}, y) = \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V_{t+1}(a_{t+1}, y) \}
\end{equation*}
subject to their budget constraint
\begin{equation*}
  c_{t} =a_{
  t}(1 +r) -  a_{t+1} + y
\end{equation*}
where $a_{t}$ are asset holdings in time t, $y$ is constant income for all periods and . Income can come either
from state pensions, defined benefit pension plans or purchased annuities.
\begin{equation*}
  u(c_{t}) = \frac{c_{t}^{1 - \sigma}}{1 - \sigma}
\end{equation*} In some specifications retirees can leave bequests, I use the bequest function from
\cite{lockwood_red_2012}.
\begin{equation*}
  B(a_{t}) = \bigl( \frac{m}{1 - m} \bigr)^{\sigma}  \frac{(\frac{m}{1 - m}c_{0} + a_{t})^{(1 - \sigma)}}{1 - \sigma}
\end{equation*}
Where $b_{t}$ is the amount left at death, $m$ is a measure of bequest motive strength and $c_{0}$ is
a minimum amount of consumption that individuals want. \textbf{Check this}

@@

I also run the regression with a variety of consumption variables.

I also simulate the expected change in consumption given forced annuitisation and no annuitisation.
The benefit of simulating an individuals decision is that we can directly compare what they would have
consumed with what they actually consumed.

### 2023-07-22 Texfiles/Notes.tex
Reorganisation of R

\subsection{ELSA notes}

>>>>

\subsection{ELSA notes}
\subsection{\textbf{20-07}}
\begin{itemize}
  \item What other ELSA variables to use?
  \item What will home ownership change?
  \item How to deal with top wealth deciles? - at the moment they are excluded because of the grid.
  \item Include other sources of income i.e. Got Dc pension
  \item Run model just with pension wealth and not financial wealth
  \item We could probably generate non annuitisation by adding some fixed cost.
  \item OK, the financial wealth variable I use is quite comprehensive including all non-housing
        wealth which is good.
  \item Deflator?
  \item

  \item Pension wealth is what I need to look at now \textbf{DONE}
  \item What
  \item Do we look at just individuals? Or just households or both?
\end{itemize}
